### 2023-01-20 my_paper/tables/tab02-observations-by-couple.tex

\begin{table}[!h]

\caption{Couples' Type \label{tab:mat2}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Couples' Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
  & \specialcell{White Husband \\ White Wife} & \specialcell{White Husband \\ Hispanic Wife} & \specialcell{Hispanic Husband \\ White Wife} & \specialcell{Hispanic Husband \\ Hispanic Wife}\\
\midrule
\textbf{Observations} & \specialcell{1,286,731\\(0.97)} & \specialcell{7,178\\(0.01)} & \specialcell{7,606\\(0.01)} & \specialcell{20,911\\(0.02)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The data is restricted to people interviewed in 1970 and 1960 and also White and married. I identify the ethnicity of a person through their place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\item[2] The table includes information on the proportion of the four types of synthetic parents that I have constructed.
\end{tablenotes}
\end{threeparttable}}
\end{table}