### 2023-01-20 my_paper/tables/tab06-regression.tex

\begin{table}[!h]

\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcc}
\toprule
  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
\midrule
$WH_{i} \times Hispanic_{i}$ & \num{-0.13}** & \num{-0.06}\\
 & (\num{0.05}) & \vphantom{2} (\num{0.05})\\
$HW_{i} \times Hispanic_{i}$ & \num{-0.16}*** & \num{-0.11}**\\
 & (\num{0.05}) & \vphantom{1} (\num{0.05})\\
$HH_{i} \times Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
 & (\num{0.05}) & \vphantom{2} (\num{0.04})\\
$WH_{i}$ & \num{0.03} & \num{0.01}\\
 & (\num{0.05}) & \vphantom{1} (\num{0.04})\\
$HW_{i}$ & \num{0.01} & \num{0.04}\\
 & (\num{0.05}) & (\num{0.05})\\
$HH_{i}$ & \num{-0.06} & \num{0.00}\\
 & (\num{0.05}) & (\num{0.04})\\
\midrule
$HW_{i} \times Hispanic_{i} - WH_{i} \times Hispanic_{i}$ & -0.03 & -0.04\\
 & (0.07) & (0.07)\\
\midrule
Observations & \num{129359} & \num{129359}\\
\textit{Controlling for:} &  & \\
Hours Worked & X & X\\
Age & X & X\\
Year FE & X & X\\
Education &  & X\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH). $Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.}
\end{tablenotes}
\end{threeparttable}}
\end{table}