### 2023-04-12 my_paper/tables/tab04-summary-stats.tex

\begin{table}[!h]

\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (i)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (ii)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (iii)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (iv)} & \specialcell{HH - WW \\ (v)} & \specialcell{HW - WH \\ (vi)}\\
\midrule
\textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
\textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\addlinespace
\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08**\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
\textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
\hspace{1em}Women’s Employment Rate & \specialcell{0.94\\(0.25)} & \specialcell{0.94\\(0.23)} & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.24)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
\addlinespace
\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
\hspace{1em}Women’s Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
\hspace{1em}Women’s Hourly Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\item[2] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents. In column one, I show the summary statistics of children of White fathers and White mothers. In column two, I present the summary statistics of children of White fathers and Hispanic mothers. In column three, I show the summary statistics of children of Hispanic fathers and White mothers. In column four, I present the summary statistics of children of Hispanic fathers and mothers.
\item[3] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
\end{tablenotes}
\end{threeparttable}}
\end{table}