### 2023-04-16 output/tables/tab03a-summary-stats.tex

\begin{table}[H]

\caption{Summary statistics of synthetic parents by couple type \label{tab:synth}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
Husband's education (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
Wife's education (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
Log Total Family Income & \specialcell{10.79\\(0.57)} & \specialcell{10.65\\(0.66)} & \specialcell{10.55\\(0.74)} & \specialcell{10.38\\(0.75)} & \specialcell{-0.41***\\(0)} & \specialcell{-0.1***\\(0.01)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}