### 2023-04-16 output/tables/tab06-regression.tex

\begin{table}[H]

\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcc}
\toprule
  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
\midrule
$HW_{i} \times Hispanic_{i}$ & \num{-0.02} & \num{-0.02}\\
 & (\num{0.08}) & (\num{0.07})\\
$Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
 & (\num{0.05}) & (\num{0.05})\\
$HW_{i}$ & \num{-0.03} & \num{0.00}\\
 & (\num{0.08}) & (\num{0.06})\\
\midrule
State FE & X & X\\
Year FE & X & X\\
\textit{Controlling for:} &  & \\
Hours Worked & X & X\\
Age & X & X\\
Education &  & X\\
Observations & \num{2423} & \num{2423}\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother, and Hispanic is an indicator variable that is equal to one if a person self-identifies as Hispanic.}
\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
\end{tablenotes}
\end{threeparttable}}
\end{table}