### 2023-04-17 output/tables/tab04-summary-stats.tex

\begin{landscape}
\begin{ThreePartTable}
\begin{TableNotes}
\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\end{TableNotes}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endfirsthead
\caption[]{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic  \textit{(continued)}}\\
\toprule
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endhead
\midrule
\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
\textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08**\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
\textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s Unemployment Rate & \specialcell{0.07\\(0.74)} & \specialcell{0.06\\(0.77)} & \specialcell{0.08\\(0.73)} & \specialcell{0.07\\(0.74)} & \specialcell{-0.00***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
\addlinespace
\hspace{1em}Women’s Unemployment Rate & \specialcell{0.06\\(0.75)} & \specialcell{0.06\\(0.23)} & \specialcell{0.07\\(0.74)} & \specialcell{0.06\\(0.76)} & \specialcell{-0.00***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\*
\end{longtable}
\end{ThreePartTable}
\end{landscape}