### 2023-05-09 output/tables/tab03a-summary-stats.tex

\begin{table}[H]

\caption{Summary statistics of synthetic parents by couple type \label{tab:synth}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{11.77\\(1.87)} & \specialcell{10.25\\(2.44)} & \specialcell{8.64\\(2.01)} & \specialcell{-4.11***\\(0.00)} & \specialcell{-1.52***\\(0.00)}\\
Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{10.40\\(2.23)} & \specialcell{11.11\\(1.75)} & \specialcell{8.49\\(1.92)} & \specialcell{-3.98***\\(0.00)} & \specialcell{0.70***\\(0.00)}\\
Total Household seducation (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{22.17\\(4.05)} & \specialcell{21.36\\(4.10)} & \specialcell{17.13\\(3.88)} & \specialcell{-8.09***\\(0.00)} & \specialcell{-0.81***\\(0.01)}\\
Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.55\\(0.26)} & \specialcell{10.46\\(0.27)} & \specialcell{10.27\\(0.21)} & \specialcell{-0.45***\\(0.00)} & \specialcell{-0.10***\\(0.00)}\\
Fertility & \specialcell{3.77\\(0.40)} & \specialcell{3.98\\(0.71)} & \specialcell{4.15\\(0.79)} & \specialcell{4.23\\(0.64)} & \specialcell{0.46***\\(0.00)} & \specialcell{0.17***\\(0.00)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}