### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats-mex.tex

\begin{table}[H]

\caption{Summary statistics of synthetic parents by couple type (Mexican Hispanics) \label{tab:synthmex}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{10.80\\(1.41)} & \specialcell{9.14\\(1.52)} & \specialcell{7.85\\(1.24)} & \specialcell{-4.90***\\(0.00)} & \specialcell{-1.65***\\(0.01)}\\
Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{9.27\\(1.71)} & \specialcell{10.46\\(1.32)} & \specialcell{7.80\\(1.36)} & \specialcell{-4.67***\\(0.00)} & \specialcell{1.19***\\(0.00)}\\
Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{20.06\\(3.09)} & \specialcell{19.60\\(2.80)} & \specialcell{15.65\\(2.60)} & \specialcell{-9.57***\\(0.00)} & \specialcell{-0.47**\\(0.01)}\\
Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.42\\(0.14)} & \specialcell{10.33\\(0.12)} & \specialcell{10.20\\(0.07)} & \specialcell{-0.52***\\(0.00)} & \specialcell{-0.09***\\(0.00)}\\
Fertility & \specialcell{3.77\\(0.40)} & \specialcell{4.26\\(0.64)} & \specialcell{4.33\\(0.64)} & \specialcell{4.46\\(0.49)} & \specialcell{0.69***\\(0.00)} & \specialcell{0.07***\\(0.00)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Mexico. A parent is White if they were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}