### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats-nonmex.tex

\begin{table}[H]

\caption{Summary statistics of synthetic parents by couple type (Non-Mexican Hispanics) \label{tab:syntnonmex}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{13.35\\(1.14)} & \specialcell{12.63\\(1.81)} & \specialcell{10.41\\(1.88)} & \specialcell{-2.34***\\(0.00)} & \specialcell{-0.72***\\(0.01)}\\
Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{12.46\\(1.22)} & \specialcell{12.46\\(1.49)} & \specialcell{10.18\\(1.74)} & \specialcell{-2.29***\\(0.00)} & \specialcell{0.00***\\(0.01)}\\
Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{25.80\\(2.18)} & \specialcell{25.08\\(3.12)} & \specialcell{20.59\\(3.50)} & \specialcell{-4.63***\\(0.00)} & \specialcell{-0.72**\\(0.01)}\\
Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.76\\(0.24)} & \specialcell{10.69\\(0.28)} & \specialcell{10.36\\(0.27)} & \specialcell{-0.36***\\(0.00)} & \specialcell{-0.07***\\(0.00)}\\
Fertility & \specialcell{3.77\\(0.40)} & \specialcell{3.52\\(0.49)} & \specialcell{3.71\\(0.80)} & \specialcell{3.85\\(0.61)} & \specialcell{0.09***\\(0.00)} & \specialcell{0.19***\\(0.00)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country other than Mexico. A parent is White if they were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}