### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03b-summary-stats.tex

%\begin{landscape}
\begin{ThreePartTable}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\caption{Summary statistics of children's outcomes using parent's place of birth \label{tab:c&p1}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endfirsthead
\caption[]{Summary statistics of children's outcomes using parent's place of birth  \textit{(continued)}}\\
\toprule
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endhead
\midrule
\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
\endfoot
\bottomrule
\multicolumn{7}{l}{\parbox{15cm}{$^1$Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes}}\\
\multicolumn{7}{l}{\parbox{15cm}{$^2$The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.}}\\
\endlastfoot
\textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men's education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.58\\(2.40)} & \specialcell{13.22\\(2.34)} & \specialcell{12.90\\(2.31)} & \specialcell{-0.92***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
\hspace{1em}Women's education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.79\\(2.44)} & \specialcell{13.42\\(2.38)} & \specialcell{13.24\\(2.39)} & \specialcell{-0.82***\\(0.01)} & \specialcell{-0.37**\\(0.02)}\\
\textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men's Unemployment Rate & \specialcell{0.04\\(0.80)} & \specialcell{0.05\\(0.77)} & \specialcell{0.07\\(0.75)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
\addlinespace
\hspace{1em}Women's Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.05\\(0.22)} & \specialcell{0.06\\(0.76)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
\hspace{1em}Men's Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.44\\(0.47)} & \specialcell{2.43\\(0.45)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0.00)} & \specialcell{-0.01**\\(0.01)}\\
\hspace{1em}Women's Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.32\\(0.46)} & \specialcell{2.29\\(0.46)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.03**\\(0.01)}\\
\hspace{1em}Men's Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.12\\(1.05)} & \specialcell{10.08\\(1.01)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
\hspace{1em}Women's Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.51\\(1.63)} & \specialcell{9.46\\(1.59)} & \specialcell{9.53\\(1.50)} & \specialcell{0.10**\\(0.02)} & \specialcell{-0.05**\\(0.05)}\\
\addlinespace
\textbf{Panel C: Children's Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men & \specialcell{0.04} & \specialcell{0.74} & \specialcell{0.83} & \specialcell{0.96} &  & \\
\hspace{1em}Women & \specialcell{0.05} & \specialcell{0.78} & \specialcell{0.82} & \specialcell{0.97} &  & \\*
\end{longtable}
\end{ThreePartTable}
%\end{landscape}