### 2024-01-15 manuscripts/economic-inquiry/manuscript_title_etc.tex
submission to EI

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 1. Define Keywords, JEL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\PAPERKEYWORDS}{\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice}
\newcommand{\PAPERJEL}{\textbf{JEL}: J71; J64; J15}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 2. Define Title
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 3. Define Authors contact information
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\AUTHORHADAH}{Hussain Hadah}
\newcommand{\AUTHORHADAHURL}{https://orcid.org/0000-0002-8705-6386}
\newcommand{\AUTHORHADAHINFO}{\href{\AUTHORHADAHURL}{\AUTHORHADAH}: Department of Economics, Tulane University, 6823 St. Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}, phone: +1-602-393-8077)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 4. Define Thanks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ACKNOWLEDGMENTS}{
I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan HADAH, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 5. Define Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\PAPERABSTRACT}{
% This study examines if Hispanic-sounding surnames lead to labor market discrimination. Analyzing wages within inter-ethnic marriages, it finds significant wage disparities disadvantaging individuals with Hispanic surnames compared to those with White surnames. Males with Hispanic fathers and White mothers earn less than their counterparts with White fathers and Hispanic mothers, primarily due to educational differences. However, this may indicate discrimination in education rather than an absence of labor market bias. Additionally, Hispanic-identifying men with Hispanic surnames face earnings gaps, largely explainable by education.
Do people with ethnically Hispanic names face discrimination in the labor market? In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames. We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities. Our findings show a significant wage difference favoring individuals with White surnames. Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels. However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation. Additionally, I explore the impact of self-identifying as Hispanic on earnings. Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities. This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.
\PAPERJEL}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 6. Define citation or availability of latest draft
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\PAPERDOIURL}{https://doi.org/10.1086/711654}
\newcommand{\PAPERINFO}{

 \url{\PAPERDOIURL}.
}