### 2024-08-28 manuscripts/07-economic-letters-v5/tab16-tab1-and-tab2.tex

\begin{table}[H]
\centering
\caption{Number of Children and Couples by Parental and Couples' Type}
\label{tab:combined}
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental/Couples' Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
 & \specialcell{White Father/ \\ White Mother \\ White Husband/ \\ White Wife \\ (WW)} 
 & \specialcell{White Father/ \\ Hispanic Mother \\ White Husband/ \\ Hispanic Wife \\ (WH)} 
 & \specialcell{Hispanic Father/ \\ White Mother \\ Hispanic Husband/ \\ White Wife \\ (HW)} 
 & \specialcell{Hispanic Father/ \\ Hispanic Mother \\ Hispanic Husband/ \\ Hispanic Wife \\ (HH)}\\
\midrule
\textbf{\specialcell{Observations \\ (Children)}} 
& \specialcell{6,421,328\\0.96} & \specialcell{39,048\\0.01} & \specialcell{51,277\\0.01} & \specialcell{179,827\\0.03} \\
\midrule
\textbf{\specialcell{Observations \\ (Couples)}} 
& \specialcell{5,141,737\\0.96} & \specialcell{33,097\\0.01} & \specialcell{37,847\\0.01} & \specialcell{119,749\\0.02} \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source (Children): Current Population Surveys (CPS) 1994-2019.
\item[2] Source (Couples): 1960-2000 Census.
\item[3] The sample includes Whites, who are married and between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}