### 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab15-emp-earn-regression.tex

\begin{table}[H]
\centering
\caption{Effect of Having Hispanic Last Name on Employment and Log Annual Earnings}
\label{tab:lastnamereg}
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
  & \specialcell{(3) \\ Log Annual \\ Earnings} & \specialcell{(4) \\ Log Annual \\ Earnings} & \specialcell{(2) \\ Unemployment} & \specialcell{(3) \\ Unemployment} \\
\midrule
$HW_{ist}$ & \num{-0.05}* & \num{-0.02} & \num{0.01}** & \num{0.01}* \\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) \\
Constant &  &  &  & \\
 &  &  &  & \\
\midrule
\textit{Controlling for:} & & & & \\
Education &  & X &  & X \\
Observations & \num{3621} & \num{3621} & \num{38090} & \num{38090} \\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to prime-age men for unemployment and full-time full-year wage and salary workers for earnings.}}
\item[4] {\setstretch{1.0}\footnotesize{I control for hours worked, age, and state-year fixed effects for earnings. I control for age, and state-year fixed effects for earnings.}}

\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered at the state level.}}
\end{tablenotes}
\end{threeparttable}
\end{table}