### 2025-01-22 manuscripts/08-JERP-v6/submit_journal-main.tex
reviwer comments

\documentclass[12pt,english]{article}

\input{preamble_package}
% Citation comes loading other packages, so that footnotes follow biblatex-chicago style
\input{preamble_citation}

% Responses to Referees
\usepackage{blindtext}
\newcommand{\aptxspc}{1.8}
\newcommand{\rrquote}{1.1}
\newcommand{\rrxspc}{1.5}

\begin{document}

% SUBMISSION COVER
\thispagestyle{empty}
\begingroup
  \doublespacing
  \centering
  \LARGE REVISE AND RESUBMIT\\[0.25em]
  \LARGE JOURNAL OF ECONOMICS, RACE, AND POLICY \\[0.25em]
  \LARGE SUBMISSION ID: 33481294-df2a-4cce-80ee-c20f2b9975ac \\[1.0em]
\endgroup
This document and the associated contents in the revise and resubmit version of manuscript 33481294-df2a-4cce-80ee-c20f2b9975ac for the Journal of Economics, Rac
e, and Policy. I include:
\begin{enumerate}
    \item Title page information
    \begin{itemize}
        \item Title and Author
        \item Abstract
        \item Keywords
        \item Acknowledgment
    \end{itemize}
    \item The manuscript
    \begin{itemize}
        \item Main manuscript text
        \item Main manuscript references
        \item Main manuscript figures
        \item Main manuscript tables
    \end{itemize}  
    \item \hyperref[r&r:responses]{Responses to the editor and referees comments}
    \item The online appendix, which is separately paged. It includes an online appendix title page, appendix A, appendix B, appendix C, and references for citations in the appendix.
\end{enumerate}
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Part I. Title Page, Authors, Abstract, etc. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in title page information, title, author information, abstract, etc are stored here
\input{manuscript_title_etc}
\pagenumbering{roman}
\setcounter{page}{1}

% Part I.a.1 Title
\section*{TITLE AND AUTHORS}
\subsection*{TITLE}
\begin{itemize}[label={}, leftmargin=*]
    \item Title: \textbf{\PAPERTITLE}
\end{itemize}
% Part I.a.2 Author info
\subsection*{AUTHORS}
\begin{itemize}[label={}, leftmargin=*]
    \item List of authors:
    \begin{enumerate}
        \item \AUTHORHADAHINFO
    \end{enumerate}
    \item Corresponding author:
    \begin{itemize}
        \item \textbf{\AUTHORHADAHINFO}
    \end{itemize}
\end{itemize}
\clearpage

% Part I.b Abstract
\doublespacing
\section*{ABSTRACT}
\PAPERABSTRACT
\clearpage

% Part I.c Keywords
\doublespacing
\section*{KEYWORDS}
\PAPERKEYWORDS
\clearpage

% statements

\doublespacing
\section*{Acknowledgement}
\ACKNOWLEDGMENTS
\clearpage

\subsection*{Data Availability Statement}
The data that support the findings of this study will be openly available to all researchers after the review process. For immediate information regarding the data and/or computer programs used for this study, please contact Hussain Hadah at hhadah@tulane.edu.

\subsection*{Funding Statement}

The author has no funding for the research, authorship, and/or publication of this article to report.

\subsection*{Conflict of Interest Disclosure}
The author declares that they have no conflicts of interest regarding the publication of this manuscript.

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Part II. Manuscript main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagenumbering{arabic}
\setcounter{page}{1}
\renewcommand*{\thefootnote}{\arabic{footnote}}
\doublespacing
\begingroup
  \centering
  \Large Manuscript Text\\[1em]
\endgroup
% Manuscript main text
\input{manuscript_main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Part III. Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begingroup
\setstretch{1.0}
\setlength\bibitemsep{5.0pt}
\printbibliography[title=References for Manuscript]
\endgroup
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Part IV. Tables and figures
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{manuscript_tab_fig}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART V. Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\begin{refsection}

% reset page numbers
\pagenumbering{arabic}
\setcounter{page}{1}
\end{refsection}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % PART VI. Responses to Editor and Referees
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{refsection}
% Reset section numbering
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}

        \section{Responses to Editors and Referee} \label{r&r:responses}
        \begin{spacing}{\rrxspc}
            I would like to thank the editor and the anonymous referees for their insightful comments, suggestions, and effort and time in reviewing this paper. I have addressed all the comments and suggestions in the revised manuscript. Below, I provide a summary of the changes made to the manuscript in response to the comments and suggestions.
        \end{spacing}

    \newpage

    \section{Responses to Referee One}
        I would like to thank referee one for the insightful comments and suggestions. Below is a detailed response to the comments and suggestions.
    \begin{spacing}{\rrquote}
    \begin{quotation}
    \textbf{R1: } 1. My first concern relates to the threats to identification. It is feasible that children born to HW couples may differ systematically to children born to WH couples, particularly with respect to unobservable characteristics important for labor market outcomes. For example, HW pairs may exhibit differing parental characteristics related to parenting styles, parental preferences for education, gender norms and beliefs. The paper would be improved by including an expanded discussion around the threat to identification. In light of these issues, the paper would be improved by providing a more thorough discussion of the relative advantages of the current approach, compared to the conventional Oaxaca-Blinder decomposition.
    \end{quotation}
    \end{spacing}

    \begin{spacing}{\rrxspc}
        I included an extensive discussion of the differences between HW and WH couples. These differences, that includes cultural differences, in couple of sections throughout the manuscript. While I acknowledge the concern, the estimation strategy rest of the fact that selection in the marriage market decreases the differences between HW and WH couples. This is the reason why comparing children of intermarried couples provides a better comparison than comparing children of endogamous marriages, i.e. marriages where both parents are either non-Hispanic White or Hispanic. I also included a discussion of the relative advantages of the current approach compared to the conventional Oaxaca-Blinder-Kitagawa decomposition. I argue that the current approach provides a more accurate estimate of gaps that are due to discrimination. In fact, the concern the reviewer raises is one of the reasons why my approach is preferred over the Oaxaca-Blinder-Kitagawa since children of endogamous marriages are more likely to have different characteristics than children of intermarried couples, including those that are unobservable. I argue, and show from the data, that children of intermarried couples are more likely to have similar characteristics than children of endogamous couples, see sections 3.1, and 4.
    \end{spacing}

    \begin{spacing}{\rrxspc}
    \begin{quotation}
        \textbf{R1: } 2. The analysis sample is restricted to individuals who self-identify as ‘White’. It is likely that this self-identification of ethnicity is endogenous to labor market outcomes. Importantly, the unobservable determinants of racial identification are potentially correlated with labor market outcomes. The paper would be improved by including a more thorough discussion of this (sample) selection issue. Related to this is the depiction of parents born in the United States as ’White’. This seems particularly over-simplifying and ignores the long history of immigration. A more accurate description would be to refer to this group as US born.
        \end{quotation}
        \end{spacing}

    \begin{spacing}{\rrxspc}
        I thank the reviewer for this comment and concern over the endogeneity of self-identification and labor market outcomes. I added some discussion over why I chose those who self-identify as White. I argue that inclusion those that are not White could contaminate the estimate of bias against Hispanics due to the racial signal. Moreover, \autocite{hadah2024effect} shows that there exists strong correlation between bias against Hispanics and self-reported Hispanic identity. Their results show that by not including Hispanics that attrit in the estimation of racial and ethnic gaps, the estimate of bias against Hispanics could be overestimated in the most biased states---more on ethnic attrition of Hispanics could be found in \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. Regarding the comment of oversimplifying and ignoring the long history of immigration in the US, the majority of immigrants from Spanish-speaking countries in the US are first- and second-generation (see \autocite{antman2020ethnic}). I also show that the probability that parents are second-generation+ immigrants from Spanish-speaking countries is very low using Census data from 1960 to 2000, see Table 5. 
    \end{spacing}

    \begin{spacing}{\rrxspc}
        \begin{quotation}
            \textbf{R1: } 3. The discussion on the construction of the ‘synthetic parents’ is somewhat brief. The paper suggests that the potential parents are matched using the birth year of the child and the parent’s place of birth to the children’s information collected in the CPS sample. While not explicitly stated in the paper, I assume that this sample of synthetic parents is used to construct mean parental education and family income at the time of the birth of the child. The process for matching the potential parents to the children is somewhat aggregated and their is likely considerable heterogeneity in the mean educational attainment of the potential parents within the set of children with the same birth year with the same parent’s country of birth. The paper could be improved by exploring the possibility of improving the quality of the match of potential parents by including further characteristics to match the potential parents to their children.
    \end{quotation}
        \end{spacing}

    \begin{spacing}{\rrxspc}
            I appreciate this insightful suggestion. I have expanded the discussion and provided a detailed example of the 'synthetic parents' construction methodology. While additional matching dimensions would be valuable, the limitations of publicly available data constrain the potential matching criteria. Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.
    \end{spacing}

    \begin{spacing}{\rrxspc}
        \begin{quotation}
            \textbf{R1: } 4. The description of the estimated model is somewhat brief. However, there are two main issues associated with statistical inference on the estimated parameters in model (1). First, there is a ‘generated regressor’ issue associated with using estimated group-level parental education. A failure to account for this sampling variation will lead to misleadingly small standard errors. Second, given this group structure for parental education, it seems reasonable to assume that the model errors are uncorrelated across clusters but correlated within (potential parents) clusters. It is well understood that failing to account for this ‘clustering’ problem can lead to misleadingly small standard errors, narrow confidence intervals, and low p-values.
        \end{quotation}
    \end{spacing}

    \begin{spacing}{\rrxspc}
        Thank you for these important methodological points. Regarding the generated regressor issue: The parental characteristics are included solely as controls and not for inference purposes. I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names. Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
    \end{spacing}

    \begin{spacing}{\rrxspc}
        \begin{quotation}
            \textbf{R1: } 5. The paper would be improved through an expanded discussion of the impact of measurement error on the reported estimates. While there is a well understood result in the measurement error literature that measures of group mean parental education will provide estimates that are more robust to the presence of measurement error in individual level measures of parental education, there is still an issue with the non-random attrition of potential parents. Specifically, not all potential parents have children in the CPS sample, nor do all children in the CPS sample have parents in the sample of potential parents. Moreover, the non-random attrition of potential parents implies that mean parental education may be systematically higher or lower than the actual parental education of the children in the CPS sample. The paper would be improved by providing an expanded discussion of the likely impacts of measurement error on the reported estimates.
        \end{quotation}
    \end{spacing}

    \begin{spacing}{\rrxspc}
        Thank you for these important methodological points. Let me clarify that the synthetic parents are constructed using the birth year of the child. The place of birth of parents---mothers and fathers---are questions that are asked of all participants in the Current Population Survey (CPS) starting 1994. Therefore, I do not use the information of parents that are in the CPS sample at the time of the survey. Thus, the concern over the attrition of parents in the sample is not applicable to my analysis. Please see the data section for a discussion of the data used in the analysis.
    \end{spacing}
    \clearpage
    \pagebreak

    \section{Responses to Referee Two}
    I would like to thank referee two for the insightful and constructive comments and suggestions. Below is a detailed response to the comments and suggestions.

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 1. The paper needs to be more clearly motivated in the introduction. There are a lot of reasons why studying discrimination by race and ethnicity is important and the paper would benefit from clearly articulating this, including discussing the implications of this discrimination. You mention economic mobility and I think there is more you can say about this. I also think clearly stating the contributions of the paper earlier on would be important.
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the important suggestion. I have expanded the introduction to better motivate the study. I have included a discussion of the implications of discrimination on economic mobility and the contributions of the paper. I have also included a discussion of the importance of studying discrimination    
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 2. I would also like to see a clearer theory and review of the literature on the topic. For example, why should we expect that discrimination will affect educational attainment? Hispanics are one of the fastest growing groups entering college, but the extent to which they complete an associate or BA degree may reflect discrimination that happens when students are in school that ultimately translates into differences in attainment. Why might we expect differences in employment and years of education?
        Why might we expect differences by gender?
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            I appreciate this suggestion. Here are my replies to the different comments.

            First, regarding including more on why discrimination could affect educational attainment. I have expanded the part of the introduction that discusses differences in education and the channels in which discrimination can affect educational attainment. The literature on discrimination and education shows that discrimination can affect educational attainment of minorities through various channels, including differences in school quality and bias from teachers, administrators, counselors, etc. These biases could lead to differences in educational attainment by preventing students access to some schools, recommendation letters, or counselors restricting access to more advanced courses, etc. Consequently, these biases could lead to differences in educational attainment. For example, having access to advanced courses in high school or better recommendation letters could affect the likelihood of attending college.

            Second, if minorities face discrimination in access to education and the labor market, then we would expect differences in employment and years of education. The literature shows that discrimination can affect labor market outcomes through various channels, including differences in access to jobs and wage differences.

            Finally, to address the comment here, and in other places, on why we might expect differences between genders. I added a discussion of why we might expect differences between men and women in the results section. I believe that showing that there might be heterogeneity in gaps between men and women in educational outcomes to be an important contribution of the paper since the literature studying gaps in earnings mainly focuses on the average gap between men. Moreover, showing that couples with a Hispanic husband and a White wife do not invest differently in their children than couples with a White husband and a Hispanic wife could be a way to test for cultural differences between the two groups. 
        \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 3. Occupational segregation can also reflect discrimination in the labor market with important implications for economic mobility and other outcomes, and it would be important to mention this in the literature review. In general, the review of the literature needs to provide more details about the different studies and how your paper contributes to that literature above and beyond using a cleaner comparison.
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for this suggestion. I have included a discussion of occupational segregation in the literature review and added how my paper contributes to the different strands of the literature.
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 4. There are places throughout the paper where additional references are needed. For example, when you state in the introduction on page 3 that discrimination can lead to lower wages, reduced opportunities, and hinder assimilation this statement needs references.
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for pointing this out. I have added references to such statements. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 5. Can you explain why you discuss assimilation and in what ways this is connected to your theory given that you are focusing on U.S. born children? If this matters because you are focusing on U.S. born Hispanic children with one foreign-born parent, then you need to clearly state this.
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the comment. I added a footnote to the introduction to address it. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 6. You also need to state clearly and early in the paper that your study focuses on children with a foreign-born parent. Children with U.S. born Hispanic parents may be different than those with foreign-born parents and you could more explicitly discuss this in the paper.
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            I appreciate this comment. I made it more clear throughout the paper that I focus on children with a foreign-born parent. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 7. Also given the differences in the characteristics of immigrants from different Spanish-speaking countries living in the U.S. (differences in socioeconomic status, education, etc.), you probably want to mention that your study is not capturing this. Further, to the extent that men and women migrate from different countries and have different pre- and post-migration characteristics then this might affect your results.

        One possible way to address this is to conduct sensitivity analyses limited to Children whose parents were likely born in Mexico or who respond themselves that they are Mexican. Given that immigrants from Mexico are the largest immigrant group from a Spanish speaking country in the U.S., this may offer an even cleaner comparison. There could still be differences in the characteristics of Mexican mothers and fathers, but you could potentially check this in the data.        
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for this suggestion. I made it clear in the paper that I do not capture differences in the characteristics of immigrants from different Spanish-speaking countries. I also added some results and discussion as a sensitivity analysis breaking down the results for Hispanic children with Mexican parents versus Hispanic children with non-Mexican parents. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 8. There are several sections that are repeated in the paper and I would suggest you streamline the text. For example, you review results twice, but this isn’t necessary.

        Similarly, I think you can more systematically organize the section describing your empirical strategy and clearly explaining your identification strategy, the concerns that it helps you overcome, your assumptions, and how you are testing whether these assumptions likely hold. Currently this is explained in multiple sections throughout the paper.        
        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            I appreciate this comment. I have streamlined the text and reorganized the paper to address these concerns. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 9. Can you provide an example that illustrates how you link individuals in the CPS to the synthetic parents?

        \end{quotation}
        \end{spacing}
            I added an example to the data section that illustrates how I link individuals in the CPS to the synthetic parents.
        \begin{spacing}{\rrxspc}

    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 10. I would like more details about the sample and the decisions you make. For example, you identify your sample as U.S. born children who identify as White in the CPS but some of these may also identify as Hispanic no? Do you restrict your sample to respondents who identify as White Hispanic or not? How many individuals in your sample identify as White non-Hispanic even if they have a parent born in a Spanish speaking country and how many do not? If someone has a Hispanic parent but do not identify as Hispanic those may be different than people with a Hispanic parent who identify as Hispanic.

        If there are people who identify as White non Hispanic in your sample even though they have a parent born in a Spanish speaking country, can you do a sensitivity analysis removing them?

        For example, on page 18 you say “I also find a significant earnings gap between those that identify as Hispanic.” Can you explain this statement? Is this related to the point I made above?

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the comment. I added more description of the sample to the data section and the reasoning of why I chose those who self-identify as White. 
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 11. You report results for men and women and, I am sorry if I missed it, but I would like to see this motivated in the paper as there are many reasons for doing the analyses separately. See my earlier point on the lit review/theory.

        \end{quotation}
        \end{spacing}
            Please see my response to the earlier comment on why we might expect differences between men and women.
        \begin{spacing}{\rrxspc}

    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 12. It would also be important to discuss effect sizes. Some of the findings seem to be small and you should discuss whether they are economically meaningful. Further, there should be more discussion about what your findings mean to understand discrimination for Hispanics and how it plays out, including by gender and in what ways they fall short in answering this question.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the suggestion. I have included a discussion of the effect sizes and the economic meaning of the results. I have also included a discussion of what the findings mean for understanding discrimination against Hispanics and how it plays out.
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 13. How are you measuring the different outcomes? I don’t believe you discuss in the paper. This may be obvious, but I think it is still important to mention for clarity.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the comment. I have included a discussion of how I measure the different outcomes in the data section.        
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 14. When you control for education in models that examine earnings, the effect becomes statistically insignificant, what happens if you include industry or occupation fixed effects? Can you do that? Then you'd be comparing people within the same industry or occupation.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            I appreciate this comment. I have added a discussion of the results when I include occupation fixed effects to the sensitivity analysis section.
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 15. The conclusion repeats much of what was said in the body of the paper, and I would like to see more discussion about the implications of your findings to understand disparities in education and labor outcomes between people with different ethnic background, and to what extent you can conclude these differences can be attributed to discrimination. It would also be important to compare your results with other literature on this topic.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you so much for this comment. I have expanded the conclusion to include more discussion about the implications of the findings and how they can be attributed to discrimination. I have also included a discussion of how my results compare to the literature on this topic.
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 16. On page 14, this sentence seems to be incomplete: “Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.”

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for pointing this out. I have corrected this sentence.
    \end{spacing}

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R2: } 17. Throughout the paper you refer to people with a parent born in a Spanish speaking country as “children who have a Spanish-sounding last name” I would encourage you to soften this language and say “who likely have” a Spanish sounding last name. First, you do not know if a person actually does have the last name of their father. Second, while it is true that Spanish sounding last names are very common in Spanish-speaking Latin American countries and among Latinos in the U.S., this may not be true across the board and many people born in Spanish speaking countries, who are Spanish speakers and Hispanic, may not have a traditionally sounding Spanish last name.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for your comment. I changed the mention of "children who have a Spanish-sounding last name" to "children who likely have a Spanish-sounding last name" throughout the paper.
    \end{spacing}
    \end{refsection}

    \clearpage
    \pagebreak

    \section{Responses to Referee three}
    I would like to thank referee three the comments. Below is a detailed response to the comments and suggestions.

    \begin{spacing}{\rrquote}
        \begin{quotation}
        \textbf{R3: } 1. As the author states on page 4, the key identifying assumption is that people born to HW parents are similar to their WH peers. Only when this assumption is fulfilled can the author attribute the estimated coefficients to evidence of discrimination. However, Table 4 clearly shows that the differences between HW and WH synthetic parents’ characteristics, such as the father’s/mother’s education and total family income, are all significantly different. More specifically, HW families exhibit lower levels of education and income compared to their WH counterparts. Based on this, it is not clear whether the estimated coefficients are indeed evidence of discrimination or they merely reflect the fact that HW children grew up in an environment with less resources/educated parents.

        The manuscript heavily adopts the methodology of Rubinstein and Brenner (2014), which examines the impacts of having a Sephardic-sounding surname on wages by comparing the Israeli-Jewish men born to Sephardic fathers and Ashkenazi mothers (SA) with those born to Ashkenazi fathers and Sephardic mothers (AS). They too find that the AS and SA families have statistically different educational and labor market outcomes. However, in their case, despite the fact that Sephardic Israelis face tougher labor market conditions, SA parents exhibit better education and labor market outcomes than AS parents. Therefore, any evidence indicating that SA offspring have worse labor market outcomes provides convincing (and potentially a lower bound) evidence of discrimination.

        \end{quotation}
        \end{spacing}

        \begin{spacing}{\rrxspc}
            Thank you for the comment. Even though the synthetic parents have different characteristics, the estimation strategy rests on the fact that selection in the marriage market decreases the differences between HW and WH couples. This is the reason why comparing children of intermarried couples provides a better comparison than comparing children of endogamous marriages, i.e. marriages where both parents are either non-Hispanic White or Hispanic. I also included a discussion of the relative advantages of the current approach compared to the conventional Oaxaca-Blinder-Kitagawa decomposition. I argue that the current approach provides a more accurate estimate of gaps that are due to discrimination. In fact, the concern the reviewer raises is one of the reasons why my approach is preferred over the Oaxaca-Blinder-Kitagawa since children of endogamous marriages are more likely to have different characteristics than children of intermarried couples, including those that are unobservable. I argue, and show from the data, that children of intermarried couples are more likely to have similar characteristics than children of endogamous couples.
    \end{spacing}

\clearpage
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART VII. Online Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{refsection}
% Reset section numbering
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}

% Title Page for Appendix
% special footnote symbols
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\begingroup
\doublespacing
\centering
\Large ONLINE APPENDIX \\[1.5em]
\LARGE \PAPERTITLE \\[0.75em]
\large
\AUTHORHADAH\footnote[1]{\AUTHORHADAHINFO}\\[1.0em]
\endgroup
\clearpage

% Online appendix
\input{manuscript_online_appendix}

\newpage
\pagebreak

% Appendix Bibliography
\begingroup
\setstretch{1.0}
\setlength\bibitemsep{0pt}
\printbibliography[title=References for Online Appendix]
\endgroup
\pagebreak
\end{refsection}

\end{document}