### 2025-06-15 manuscripts/10-JERP-v3/tab10-regression-weekearn-cuba.tex

\begin{table}[H]
\centering\centering
\caption{Effect of Having Hispanic Last Name: Hispanics with Cuban Ancestry  \label{tab:lastnamereg-weekearm-cub}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lccc}
\toprule
  & \specialcell{(1) \\ Log weekly \\ earnings} & \specialcell{(2) \\ Log weekly \\ earnings} & \specialcell{(3) \\  Log weekly \\ earnings}\\
\midrule
$HW_{ist}$ & \num{-0.09}*** & \num{-0.06} & \num{-0.01}\\
 & (\num{0.03}) & (\num{0.06}) & (\num{0.04})\\
Constant & \num{6.73}*** &  & \\
 & (\num{0.04}) &  & \\
\midrule
\textit{Controlling for:} &  &  & \\
State FE &  & X & X\\
Year FE &  & X & X\\
State-Year FE &  & X & X\\
Age &  & X & X\\
Education &  &  & X\\
Observations & \num{1318} & \num{1318} & \num{1318}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}) where the dependent variable is log weekly earnings.}}
\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time and are wage and salary workers.}}
\item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects. Column two has the results after controlling for education.}}
\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
\end{tablenotes}
\end{threeparttable}
\end{table}